---
title: "Exploratory Analysis - Kitchen Stoves"
author: "JJ Espinoza"
date: "Sunday, August 23, 2015"
output: word_document
---

The following exploratory analysis examines data on product characteristics of kitchen stoves. 


**Stove Requirements**

Stainless steel


Dimensions of the stove: Depth = 25in, Width = 30in, Height 45in

**Key Findings (statistically significant):**

- The average price of a stainless steel stove with gas burners is $1,500

- The dimensions of the stove we need is common

- There is a strong correlation between price and weight; likely due to better materials

- The capacity of the stove is also a key driver, specifically the capacity of the oven

- The effectiveness of the stove is important but mainly driven by the heating capacity of the oven


##Importing the data.

```{r echo=TRUE}
data <- read.csv("C:/Users/ESPIJ090.WDW/Home - Kitchen Stoves/data/Kitchen Stove.csv")
```



###We want to limit our search to stainless steel stoves only.

```{r echo=TRUE}
stainless <- data[which(data$Color=="Stainless steel" & data$CookingSurface=="Gas: sealed burners"),]
```

##Data Visualization and Analysis

###What should we expect to pay for a stove?

```{r echo=TRUE}
hist(stainless$Price, col = "red", main="Stainless Stove Prices", xlab = "Price($)")

summary(stainless$Price)
```

###What are the common dimmensions of stoves in relation to what we are looking for?

We are looking for a stove that is no more than 30in.

```{r echo=TRUE}
hist(stainless$Width, col = "blue", main="Width of Stainless Stoves ", xlab = "Width (in)")
summary(stainless$Width)
```

We are looking for a stove that is close to 45 in height.
```{r echo=TRUE}
hist(stainless$Height, col = "blue", main="Height of Stainless Stoves ", xlab = "Height (in)")
summary(stainless$Height)
```

We are looking for a stove that is close to 25 in depth.
```{r echo=TRUE}
hist(stainless$Depth, col = "blue", main="Depth of Stainless Stoves ", xlab = "Depth (in)")
summary(stainless$Depth)
```

###What are the tradeoffs between stove dimensions and price? 

The correlation matrix shows a strong correlation between price and weight.
```{r echo=TRUE}
library(car)
scatterplot.matrix(~Price+Depth+Height+Width+Weight , data=stainless, main="Price Correlated with Dimensions of Stove")
```

Taking a closer look one can see the clear relationship.

```{r echo=TRUE}
plot(y = stainless$Price, x = stainless$Weight, pch = 19, col = "red",main = "Correlation between Weight and Price", xlab="Weight (lbs)", ylab = "Price ($)")

abline(lm(stainless$Price~stainless$Weight), col="red") # regression line (y~x) 
```

Regression estimate of relationship between price and weight.
```{r echo=TRUE}
summary(lm(Price~Weight, data = stainless))
```

###What other features are correlated with price?

```{r echo=TRUE}
scatterplot.matrix(~Price+ MainOvenCapacityCuFt+ NumberofElementsBurners  , data=stainless, main="Price Correlated with Dimensions of Stove - Plot1")
```


```{r echo=TRUE}
scatterplot.matrix(~Price + MainOvenNumberofRackPositions + NumberofOvenRacks , data=stainless, main="Price Correlated with Dimensions of Stove - Plot2")
```

Taking a closer look at the correlations betwen Number of Oven Racks and Price we see a negative relationship.

```{r echo=TRUE}
plot(y = stainless$Price, x = stainless$NumberofOvenRacks, pch = 19, col = "red",main = "Correlation between Number of Oven Racks and Price", xlab="Number of Oven Racks", ylab = "Price ($)")

abline(lm(stainless$Price~stainless$NumberofOvenRacks), col="red") # regression line (y~x) 
```

Regression estimate of relationship between price and number of oven racks.
```{r echo=TRUE}
summary(lm(Price~NumberofOvenRacks, data = stainless))
```

###How does the effectiveness and efficiency correlate with price?

The scatterplot below shows that the higher the BTUs the lower the price, which seems counterintuitive.

```{r echo=TRUE}
scatterplot.matrix(~Price + BroilerBurnerBTUs +  MainOvenBakeBurnerBTUs , data=stainless, main="Price Correlated with Effectiveness of Oven")
```



##Regression Analysis
```{r echo=TRUE}
model <- lm(Price ~ Weight + MainOvenCapacityCuFt + NumberofOvenRacks + BroilerBurnerBTUs +  MainOvenBakeBurnerBTUs , data=stainless )
```

```{r echo=TRUE}
summary(model)
```

```{r echo=TRUE}
plot(model)
```

###How good does the regression model fit the data?

```{r echo=TRUE}
plot(x = predict(model, stainless), y = stainless$Price, main = "Model Predictions vs. Actual Prices", xlab = "Predicted Price", ylab = "Actual Price")
abline(0, 1)
```

##The model accuracy is: 

```{r echo=TRUE}
mean(abs((stainless$Price - predict(model, stainless))/stainless$Price), na.rm=TRUE)

```

###Appendix - Summary Statistics

Summary statistics

```{r echo=TRUE}
summary(stainless)
```


